### 2018-04-03 paper/privacy.tex
Pontless commit

\section{Privacy}%
\label{Privacy}

We care about
\dots