### 2018-04-17 paper/primitives.tex
Adds primitives section

\section{Primitives}%
\label{Primitives}

The algorithm (\cref{Algorithm}) makes use of several primitives, which we 
describe here.
There is a local prediction oracle (described in \cref{LocalPrediction}) which 
is used by a device to predict its own online behaviour.

\subsection{Local prediction oracle}%
\label{LocalPrediction}

\NewScheme{\LP}{P_L}
\NewAlgorithm{\LNext}{Next}
\NewAlgorithm{\LProb}{Prob}

We have a local prediction oracle, \(\LP\), which provides the following two 
algorithms:
\begin{description}
  \item[\(\LNext\)] Test \dots
  \item[\(\LProb\)] \dots
\end{description}
We will denote these as \(\LP[\LNext]\) and \(\LP[\LProb]\).

\subsection{Global prediction oracle}