### 2018-04-25 paper/building-blocks.tex
Adds building blocks section

\section{Building blocks}%
\label{BuildingBlocks}

\dots