### 2018-04-28 paper/file-exchange.tex
added files...

\subsection{The file exchange}

\commentDaniel{%
  We must do the scheduling: \ie divide the file \(f\) into \(f_1, \dotsc, f_n\) 
  and how the size of \(f_i\) relates to transmission time and size of the 
  timeslot (for the validity of predictions).
  After that we just do \(\SPOR[\Send]\) on each \(f_i\).%
}

\subsubsection{File exchange process}% (fold)
\label{ssub:file_exchange_process}

Firstly, Alice and Bob must exchange information out-of-band: an encryption key $ek$ to cipher the file, and an ID $k$ in the DHT pointing to information on how to reach Alice.