### 2018-04-29 paper/conclusion.tex

\section{Conclusion and Future Work}%
\label{conclusion}

We presented \(\SPOR\) that, in analogy to its biological
counterparts, enables dissemination under unfavorable conditions. More
specifially, the deliberate dissemination of files without using
storage on third-party devices, be they peers or servers in the
cloud. This constraint results in the unfavorable conditions of
reduced a-priori availability and therefore unreliable nodes for
routing. We build resilience to these conditions by using predictive
modeling for user behavior and the ensuing availability of their
devices. From our experiments, we find that even suboptimal prediction
already results in successful file transfer with high probability,
though often at low speed. More in-depth knowledge of the users'
behavior would allow us to predict device connectivity farther into
the future and thus likely improve the throughput. From a security and
privacy perspective, we were guided by the properties of
TOR~\cite{Tor}, which we modified to a stateless protocol to
accommodate unreliable nodes. While we took the first steps toward
analyzing the consequences of the modifications on security and
privacy, more work is needed to provably ensure that the added feature
of resilience despite unreliable nodes (supported by peer sampling)
and the side effect of enabling more local TOR routing (for example,
keeping routes inside a country) do not compromise security. On a
related note, it would be interesting to adapt the
Sphinx~\cite{Sphinx} cryptographic packet format to the routing of
\(\SPOR\) as a step toward provable security.