### 2018-07-19 paper/file-exchange-design.tex
Finalizes editing of design

\subsection{Transferring files over \acs{SPORES}}%
\label{sec:file_exchange}

We will now show how to build a file transferring service on top of 
\ac{SPORES}.