### 2018-08-06 paper/system-model.tex

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex -halt-on-error p2p-private-cloud.tex > /dev/null" ispell-dictionary: "american" -*-
%!TEX root = p2p-private-cloud.tex

\section{System model, adversary model and desired privacy properties}%
\label{system-model}

\name provides a private and fully decentralized file transfer service by exploiting 
%% is a distributed system made of
users' devices.  We assume any two devices can communicate with one
another as soon as one knows the address of the other (\eg using the
Internet, possibly extended with NAT traversal techniques). The
devices owned by the same user (this user's \squad) know of each
other. Other users' devices can help transfer the file from one user
to another but are not needed for storage.  We also assume that users
can exchange some limited amounts of data out-of-band, \ie through
some other channel than the network just mentioned (\eg using \ac{NFC}
or QR codes).
%system model: explain e-squad, contrast with cloud

%adversary model: remove provider, compensate, results in same as in sphinxOR/Tor?  friends?

%properties: cam/lys: correctness, integrity, security (wrap, indistinguishability) - do not want friends to learn 1) which devices are ours 2)  their availability (assumption: RPS, not lookup)

In the following, we first describe how file transfer typically occurs
when users rely on a cloud storage provider. We use this simple use
case to highlight some of the desirable privacy properties of a file
transfer service, before mapping them into our fully decentralized
(a.k.a. P2P\footnote{We use the terms \emph{fully decentralized} and
  \emph{P2P} interchangeably, and preferentially the latter for brevity.})
context.  We finally consider the additional privacy properties
required to protect users from inferences such as by network traffic
analysis and state our adversary model.

\subsection{File transfer via a cloud storage provider}

When sharing a file with Bob through a cloud provider, Alice only
needs to upload this file from one of her devices to the cloud, and
share a link to this file with Bob. Bob in turn can download this file
through the transmitted link onto one of his devices.

The use of the cloud provider, as a third party,
provides two main privacy properties, which we should preserve when
moving to a \ac{P2P}~\cite{DevilInMetadata} setting:
\begin{itemize}
\item Alice and Bob are the only users who learn 
about Alice sharing a file with Bob.
\item The composition and behavior of Alice and Bob's \squad remain
  hidden to each other, and to other users. I.e. Alice does not learn
  which devices Bob owns or uses, let alone how Bob uses them, and
  reciprocally.
\end{itemize} 
The cloud provider, however, has this information and any other
stemming from the users' interaction with the service, and can draw
inferences from metadata even when the data itself is encrypted, and
thus poses a threat to the users' privacy.

\subsection{Moving to a P2P environment}
A private P2P file transfer solution should retain the privacy that
Alice and Bob already enjoy when a cloud provider functions as a
gatekeeper towards each other and other users.  Eschewing a cloud
provider, as we propose to do, removes the privacy risks associated
with a central brokering entity, but also removes the gatekeeping a
cloud provider provides, and exposes users to friendly
surveillance~\cite{FriendlySurveillance}, in an honest-but-curious
setting, by other peers (other \name users including Alice and
Bob). In particular other users might collude to observe some of the
traffic, and gain knowledge regarding Alice's and Bob's interactions,
the devices they own, and how they use them.

\subsection{The privacy properties and adversary model of \name}

Friendly surveillance can be mitigated by ensuring security properties 
typically associated with onion routing, as defined by Camenisch and Lysianskay~\cite{CLOnionRouting}: 
% We combine the requirements from
% the \ac{P2P} setting with those from onion routing for anonymity.
% Camenisch and Lysianskaya \cite{CLOnionRouting} define security properties 
% for onion routing:
correctness, integrity, security (wrap resistance,
indistinguishability). These properties cover the concerns about
revealing to others that Alice shares a file with Bob in the first
place. In \name, we additionally derive the property of
device-unlinkability from the \ac{P2P} requirements, meaning that
using \name does not reveal the \squad a device belongs to nor its
owner. Thus, no information on the user's behavior (e.g. being online)
can be infered through \name.  \commentFT{Removed, as did not fit with
  the flow, but probably needed somewhere: `We give the definitions
  for the security and privacy properties of \name in
  Section~\ref{security-discussion}.'} \sonja{better here. Need to define
  device unlinkability, the rest is covered by Camenisch.}

\paragraph*{Adversary model} To assess our approach, we use the
adversary model assumed by Tor~\cite{Tor} and Sphinx~\cite{Sphinx}:
the adversary has complete control over some part of the network as
well as some of the nodes. We do not consider a passive global network
adversary, against which it is notoriously difficult to achieve
security~\cite{SystemsForAnonymousCommunication}.