### 2018-08-06 paper/security-discussion.tex
related work: p2p storage, i2p, uncommented availability modeling

\section{Security Discussion}%
\label{security-discussion}

%\sonja{will add some framing about us composing and modifying parts that
%  have security proofs for their original and individual versions and needing to evaluate the resulting s%chemes}

\commentDaniel{Predecessor attacks: random walk protocols has the problem that 
  the adversary can see the initiator (statistically) more often than the other 
  nodes, and can thus identify the 
  source~\cite{RoutingSurveyAnonymousProtocols}.
  Our protocol is a \enquote{bounded} random walk, \ie within limits.}

We base our solution on Sphinx~\cite{Sphinx}, which, when used for
onion routing, has been proved to satisfy the properties given
in~\cite{CLOnionRouting} that were shown to result in a UC-secure
onion-routing scheme. We thus primarily need to ensure that
\begin{enumerate*} \item our extension to Sphinxes preserves these
  properties, \item the availability information introduced and
  disseminated to deal with churn satisfies the device-unlinkability
  property we defined in~\cref{system-model}, and that \item the
  resulting node selection does not introduce any advantage for the
  adversary. \end{enumerate*} In terms of general system robustness, a
stateless routing protocol needs to handle packet re-ordering and
network failures (links or nodes).

\paragraph*{Security of Sphinxes}

There are two changes that we have made that must be analysed.
We changed the key agreement from a pure \ac{DHKE}, \ie based on the \ac{DH} 
assumption, to one containing an indirection.
We also changed from the classical \ac{OR} of a fixed path with one node per 
layer to \ac{POR} with several nodes per layer.

The confidentiality of the key previously rested entirely on the \ac{DH} 
assumption (\(\alpha\) in the header).
Now it also rests on the security of the \ac{PRP} \(\prp'\) (\(C\) in the 
header).
The integrity is still reduced to the security of the \ac{MAC} \(\mac\) 
(\(\gamma\) in the header).

The second change is that we use several nodes per layer as opposed to only 
one.
The proofs of Sphinx rest on the results of \textcite{CLOnionRouting}, which 
are results for ordinary \ac{OR} with only one node per layer.
\sonja{Daniel is working on sketching a proof/rephrasing the following.}
As long as we output all the nodes in the next layer, the proofs should not 
change.
We note that this is actually quite natural as a malicious adversary can drop 
any incoming traffic to a router and wait for the timeout and retransmission to 
the next node.
As such, the adversary can learn all nodes in the layer.

With these arguments, we believe that Sphinxes inherits the \ac{UC} security 
properties of Sphinx, however, this should be formally analysed.

\paragraph*{Security of \acs*{SPOR}, node selection}

\sonja{will motivate using the survey/taxonomy}
The onion route provides unlinkability between Alice and Bob as long as the 
message passes at least one honest node on the route.
This means that any adversary, say Eve,  must ensure that the message passes 
through an adversarial node on each layer.
The probability of successful attack is proportional to Eve's size in the 
network, \ie \(f_A = \frac{|D_A|}{|D|}\).
\commentDaniel{Check that we still use the \(D_i\) notation for the devices in 
  \(i\)'s squad.}
If we assume that \(\GetRandomNode\) returns nodes with close to the uniform 
distribution, then the \(\nu w_L\) nodes selected will contain \(f_A \nu w_L\) 
nodes controlled by Eve.
If \(f_A \nu w_L < \nu\), then Eve cannot be present on every layer of the 
route.
Considering how \ac{SPOR} distributed the \(\nu w_L\) selected nodes, Eve's 
best strategy is to have high-availability nodes --- assume that 
\(\avail_{D_E}(t) = 1\) for all Eve's devices \(D_E\).
This way, Eve's devices will be distributed evenly across all layers.
To succeed, Eve must be chosen in the first layer, the probability of which 
will be \(\frac{f_A}{w_L}\).
If one of her devices is chosen in the first layer, then she can choose one of 
her own devices in the next layer.
Since she is present on every layer, she can ensure she relays the message all 
the way to Bob. \sonja{compare the chance of being present on every
  layer AND chosen in the first layer to regular onion routing where she
  needs to be the (sole) node on each layer. Are we doing
  better/worse/the same?}

However, Eve can only be certain that she has relayed the message from Alice to 
Bob if \(\nu = r\), \ie she receives the message from Alice, relays it \(r-1\) 
times and lastly sends it to Bob.
If \(\nu < r\) though, then, due to the indistinguishability and unlinkability 
properties of Sphinxes (inherited from Sphinx), Eve cannot know for certain if 
Alice is the first node and Bob is the last node. \sonja{but what can
  she learn/guess?}

Finally, we must point out that the \(\GetRandomNode\) algorithm must be 
privacy preserving.
Consider the following scenario.
Alice requests a set of peers, say \(\{d_1, \dotsc, d_n\}\).
If \(d_i\) interacts in each sampling operation resulting in \(d_i\) being in 
the sample, then \(d_i\) can generate a new public key \(\pk_d^{(j)}\) for each 
sample \(j\).
This means that, if Alice sampled \(d_i\), she knows the public key 
\(\pk_i^{(j)}\), and if \(d_i\) knows it is included in Alice's sample, then it 
will also know a route belongs to Alice whenever \(\pk_i^{(j)}\) is used to 
encrypt the header.
(This follows from the fact that Sphinxes is source-routed.)
\Ie every node on Alice's route would know that Alice is either the source or 
the destination.

\paragraph*{Robustness}

Stateless routing can result in re-ordered chunks or packets, and
chunks received on different devices of a user's squad. From Sphinx,
we inherit IDs for packets that we we include in the out-of-band
metadata on the file to be exchanged, as an ordered list, that is used
to re-order chunks if need be. The squad overlay handles inter-squad
device coordination. Liveness despite the possibility of squad devices
going offline during transfer or other node or link failures is
ensured by acknowledgments and timeouts as detailed in
~\cref{sec:squad_overlay}.

\paragraph*{Extended adversary model}%
\label{security-limitations}

A global, passive, network-observing adversary can of course do
statistical disclosure attacks to eventually learn who communicates
with whom~\cite{StatisticalDisclosureAttacks}. \sonja{Daniel: the
  reference seems to be missing}  There are also denial
of service attacks that can lead to the adversary learning more from
the system~\cite{DenialOfSecurity}.  As mentioned in
\cref{system-model}, we share the adversary model with Tor and Sphinx
and thus exclude a global passive adversary. Our focus is on
protecting Alice and Bob from each other and from others who also
participate as nodes in the network, not on protecting them from
surveillance. Nevertheless, we offer some reflections on such an
adversary. We hypothesize that our stateless and redundant design
might yield some advantages over traditional onion routing. For
example, no control messages are sent back to the source if the
adversary has introduced an error.  \sonja{``But these advantages will
  only postpone the success of the attack, eventually even our
  protocol must resend.'' why?}  \commentDaniel{We should look into
  this denial of security.} We would also like to point out that
vulnerability to a global passive attacker is not inherent in Sphinx
or Sphinxes, but can be caused by the way they are applied, \ie the
traffic sent. If performance and resource consumption were not an
issue, batched and well-timed message or file transfer by all nodes
would render traffic analysis pointless.