### 2018-08-07 paper/conclusion.tex
typo in intro, started conclusion edits, more needed

\section{Conclusion and Future Work}%
\label{conclusion}

We presented \name that, in analogy to its biological
counterparts, enables dissemination under unfavorable conditions. More
specifially, the deliberate dissemination of files without using
storage on third-party devices, be they peers or servers in the
cloud. This constraint results in the unfavorable conditions of
reduced a-priori availability and therefore unreliable nodes for
routing. We build resilience to these conditions by using predictive
modeling for user behavior and the ensuing availability of their
devices. From our experiments, we find that even suboptimal prediction
already results in successful file transfer with high probability,
though often at low speed. More in-depth knowledge of the users'
behavior would allow us to predict device connectivity farther into
the future and thus likely improve the throughput. From a security and
privacy perspective, we were guided by the properties for secure onion
routing schemes defined by~\cite{CLOnionRouting} and added a property
for device-unlinkability to prevent inferences from availiability data.
% TOR~\cite{Tor}, which we used to provide a modicum of privacy for this
% first investigation of feasibility of file-sharing between two users'
% squads of devices and modified to a stateless protocol to
% accommodate unreliable nodes. While we took the first steps toward
% analyzing the consequences of the modifications on security and
% privacy, more work is needed to provably ensure that the added feature
% of resilience despite unreliable nodes (supported by peer sampling)
% does 
% %and the side effect of enabling more local TOR routing (for example,
% %keeping routes inside a country) do 
% not compromise security. Another
% consideration is the nature of the traffic in this scenario and the
% limits of protection against statistical inference.

Although we presented a stateless solution, mainly to handle churn in
a file transfer scenario, \name does not preclude a stateful alternative. 
Other applications such as web browsing, interaction and
collaboration or other kind of hidden services might benefit from a stateful, circuit-based approach
like with traditional onion routing. Prediction of device availability
could support different goals, for example time of day for delayed
file transfer for resource optimization.

%
% \commentAL{To rewrite as real paragraphs} Future work and prospects:
% \begin{itemize}
% 	\item More in-depth knowledge of the user's behavior would allow to predict devices connections farther in the future;
% 	\item Changing devices' asymmetric keys regularly would increase security.
% 	Updating keys is easy in \name given the RPS that propagates new information to all nodes.
% 	However, more reflection would be needed to avoid key updates breaking ongoing communications passing through the devices in question;
% 	\item \name is not resilient to Byzantine nodes for now: malicious users could potentially make users pick Predictive Onion Routes that pass through them with a high probability, thus allowing the attacker to learn the file sender and receiver. To circumvent this, we would need to employ a secure Random Peer Sampling \cite{Jesi_Montresor_van_Steen_2010}, and prevent devices to lie on their probability of being online.

% \end{itemize}