### 2018-08-08 paper/abstract.tex

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex p2p-private-cloud.tex" ispell-dictionary: "american" -*-
% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?
%
%abstract
%problem/topic
This paper proposes SP$^2$ORES, a novel private file transfer service that avoids any third-party storage or centralized bootstrapping service.
%research gap
Our solution, which adopts a fully decentralized design, overcomes the limitations of cloud file-sharing systems such as Dropbox, which require trust in the service provider and can be monitored, even when their data is encrypted. Our proposal also goes beyond existing private peer-to-peer (P2P) solutions that must rely on storage replication for efficiency or a fixed set of trusted servers for bootstrapping purposes, thus limiting their level of protection. % We are interested in minimizing the trust needed in other systems and preserving the users' privacy in terms of data and metadata, what they share with whom and when.
%why now
% Nowadays, many people use multiple devices and it has thus become possible to benefit from P2P technology for one's own devices alone, thereby reducing one's exposure to security and privacy threats from third parties.
%approach
In this paper, we demonstrate that users do not necessarily have to choose between efficiency and strong privacy when sharing data with others. % that is  both efficient and strongly private is possible.
\commentFT{Trying to highlight POR, as likely to catch people's attention, and contribute to the paper's long-term interest.} At the core of our proposal lies a novel stateless probabilistic onion routing protocol called SP$^2$OR, which exploits decentralized peer sampling, a probabilistic model of node availability, and a novel set-based onion routing header format (dubbed SphinxES). \commentFT{Sharing node availability information freely could potentially disclose private information. We might want to discuss this risk somewhere in the paper.}
% Our proposal combines onion routing with probabilistic node selection, informed by peer sampling and Hidden Markov Model predictions for node availability.
%findings/contributions:
The resulting solution is fully decentralized, provides sender-receiver anonymity, and can deliver competitive performance in spite of unreliable nodes. %% to protect user privacy in decentralized file sharing.
% We show that even a relatively naive prediction model suffices to share files despite the unreliable availability of individual nodes. %% The same model can potentially be applied for better-informed churn handling in other p2p systems. In the case of onion routing systems for anonymity,
\commentFT{Using transient to avoid repeating unreliable.}
This design makes it possible to exploit transient nodes, without dedicated relays, %take advantage of locality, 
neither of which is currently the case in Tor.
\sonja{say something about security requirements? About sphinxes?} \commentFT{Added a mention to SphinxES, not sure if we can discuss security requirements without making the abstract overly long.}