### 2019-02-01 spores-protocol-paper/simpleSPOR.tex
Adds copy of paper to spores-protocol-paper

\onecolumn
\section{Simple SPOR, differences to TOR}
\begin{verbatim}
Names: source S, destination D, intermediate node (set) I, rendezvous point R, file f, secret symmetric key k_s, threshold t,  a node's public key for ANOBE k_p_j

Simple \ac{Tor}:
1) S: get I from directory
2) set up circuit (via key agreements with each I_i)
3) send/receive in circuit via onion (constructed by encrypting packet for each layer, next hop outermost)

Simple SPOR:

out of band:
1) S tells D: want to send f, metadata(f)
2) D tells S: here's the onion_D(R_D) (path from R_D to D)
Onion(R_D, Onion(I_1, Onion(I_2, Onion(D, payload))))
3) agreement on k_s to encrypt f

To construct an onion:

1)  get I from random peer sampling (request until P(all nodes in set I_i down)<t)
2)  get keys for all nodes in I
3) for each I_i, encrypt for all nodes in I_i, using k_p_j
3) for each I_i, choose 1 node (at random) to send to

File sharing:
1) out of band communication
2) S creates onion for path from R_S to S:
Onion(R_S, Onion(I_1, Onion(I_2, Onion(S, payload))))

3) S creates onion for path to R_D, thus path from S to D

Onion(S, Onion(I_1, Onion(I_2, Onion(R_D, Onion(I_1, Onion(I_2, Onion(D, payload)))))))

4) S uses 3) to send payload 2) (path R_S to S) to D

5) S sends enc_k_s(f) to D, as payload in 3)

6) D creates onion for path from D to R_S, thus path from D to S

Onion(D, Onion(I_1, Onion(I_2, Onion(R_S, Onion(I_1, Onion(I_2, Onion(S, payload)))))))

7) D sends acknowledgment to S along (as payload in) 5) upon receipt of f

Web browsing:

1) last node before web server W (exit node) is treated as R_S

2) S creates onion for path from exit to S
Onion(R_S, Onion(I_1, Onion(I_2, Onion(S, payload))))

3) S creates onion for path from S to W via exit
Onion(S, Onion(I_1, Onion(I_2, Onion(R_S, Onion(W, payload)))))

4) S sends 2) along (as payload in) 3), with http request for server

Note: need to figure out how \ac{Tor} sends both return onion and request, then do the same here.

5) R_S forwards http request to W, via http(s)/TCP

6) W responds to R_S

7) R_S forwards response to S as payload to 2). (R_S keeps 2) as W is not part of SPOR)

Assumptions/questions:
- choice of t, based on what? Will be informed by performance analysis.

- set of peers for RPS, all peers in the world or possible restrictions. Restrictions possible by FilteredView but loss of security (more predictable, vulnerable to Sibyl)

- RPS gives ID(=address?), public key (ANOBE), probability of availability. Note: take into account Sibyl attacks. Probility of availability currently average, but could be different [1], etc. Problem with public key by RPS is what authentication we can achieve. \Eg 
can Mallory change the public key corresponding to some other node's 
address and thus man-in-the-middle that node? I think this is also 
reduced to the RPS we use.

- metadata(f) is size, digest. Note: possibly only digest.[2]

- only one ack upon whole file receipt, time-outs for retry at S and D. [2]

- time-outs derived from file size, availability? Adrien: This is a problem I want to address. When does S starts resending chunks that it already shared? This has a great impact on the network cost of the overall file exchange.

- TCP between hops (S to I_1 ... to D). Reason: joint probability for loss too high for UDP and maybe we want the chunk size larger than UDP can handle in one package.

- web browsing not stateless at exit node, even possible

- ANOBE encrypts for all, non-recipients get nonsense. But here we only encrypt for set of nodes in I_i. Note: current plan to use \ac{Sphinx} and modify it to allow sets of nodes (encryption from ANOBE)

Comparison between \ac{Tor} and SPOR:
- SPOR file sharing similar to \ac{Tor} hidden service
- \ac{Tor} web browsing similar to Spor
- \ac{Tor} stateful (circuits), SPOR stateless
- SPOR layers are sets of nodes, \ac{Tor} layers are single nodes.
- SPOR allows for parallelism (bottlenecks only at S and D). Whereas \ac{Tor} bottlenecks at the node in the circuit with the lowest 
bandwidth.
- \ac{Tor} infrastracture, SPOR decentralized
- SPOR allows any node intor relay directory. Anyone can enter. Then they receive different ranks based on "good behaviour". They also monitor the nodes and kick out suspicious nodes.
- SPOR chooses routes informed by availability prediction to minimize churn (prevention)
- SPOR can do local repairs (future work?) to cope with churn (detection, response, recovery). Already now: if node not available, choose another from the set to send to. Problem in current version if node is available but then crashes. [3]
- SPOR can limit location, e.g. within a country? Would help with availability and uniform distribution of chosen nodes again due to lack of a-priori differences in availability in same time zone.
- \ac{Tor} sender anonymity, SPOR file sharing sender-receiver anonymity (ID of device, not user as there's out-of-band communication), but the more fair analogy is probably with hidden services, where there should be sender-receiver anonymity also in \ac{Tor}. Only sender anonymity when SPOR is used for web browsing.

[1] Adrien: Using a vector introduces discrete time, which I would prefer to avoid.
And our prediction model won't produce significantly different results
for P(online at t+1) and P(online at T+infinity). I was thus thinking of
using only P(online at t+1) as a general "P(online in the future)"
predictor. A better prediction model would be future work (not our
point: just plug your own prediction model).

[2]
Daniel: We could let the digest be the digest of each chunk. Then we get the ordering of the chunks as well. (\Ie the metadata is equivalent to a torrent file.) Then the ack could be a bitmap of which chunks are received. This is good for efficiency but opens up for attacks. However, the damage of this is not as bad as for mixes (or \ac{Tor}), since the retransmission will *probably* not take the same route.
Adrien: We take inspiration from BitTorrent. They chunk files in fixed-size
pieces, make a SHA1 hash of each piece, and a hash of the previous
hashes. This is all we need to share. We can even share it in-band,
during your point 4) in file-sharing. Thus, we only need to exchange a
file identifier out of band.

[3] from Daniel's mail. (remark by Sonja: I don't know what the original v. current versions are)
Our local repairs are more limited when we have onion_D for acks. Then 
acks must be handled on a S--D basis.

Consider node I_i successfully transfers the chunk to I_{i+1}. Now 
I_{i+1} crashes before transferring the chunk to I_{i+2}. I_i cannot 
detect this, it must be detected by and solved by S and D.

With the original design I_i would detect this and resend the chunk to 
I_{i+1}'. In this case, I_i would also resend the chunk if I_{i+1} 
crashed after *successfully* transferring the chunk to I_{i+2}. So D 
would end up with receiving two copies of the chunk.

I still think that the original design is better from a privacy 
perspective as it hides S a bit more. The new design makes S and D more 
visible as any retransmission will force them to take action.

Adrien: We do handle repair. If I_{i} is down, there was a t (parameter)
probability that it would happen. This shouldn't last long, since t lies
around 10^-3. We don't change the route once it's created; it is crafted
for reliability from the ground up.

\end{verbatim}