### 2020-03-23 paper/spores.tex
Changes to plain article format

\documentclass{article}
\input{preamble.tex}
\usepackage{authblk}

\begin{document}
\title{%
  \acs*{SPORES}: \acl*{SPORES}%
  %\name: Stateless Predictive Onion Routing for E-Squads%
  %% From the IEEE template
  % \thanks{Identify applicable funding agency here. If none, delete this.}
  % \thanks{%
  %   Daniel Bosk and Sonja Buchegger were funded by
  %   the Swedish Foundation for Strategic Research grant SSF FFL09-0086.
  %   Research done while Daniel Bosk visited the WIDE team at INRIA/IRISA Rennes, 
  %   France.
  % }
}

\author[1]{Daniel Bosk\thanks{%
  Research done while the author visited the WIDE team at INRIA/IRISA Rennes, 
  France.
}}
\affil[1]{KTH Royal Institute of Technology\\\email{daniel@bosk.se}}
\author[1]{Sonja Buchegger}
\author[2]{Adrien Luxey}
\affil[2]{Univ.\ de Rennes, INRIA, CNRS, IRISA\\\email{Adrien.Luxey@irisa.fr}}
\author[2]{Yérom-David Bromberg}
\author[2]{François Taïani}

\mode<article>{\maketitle}
\mode<presentation>{%
  \begin{frame}
    \maketitle
  \end{frame}
}

\begin{abstract}
  \input{abstract.tex}
\end{abstract}

\input{contents.tex}
\end{document}