### 2020-03-23 paper/evaluation.tex
Removes performance eval (heavy on prediction)

%!TEX root = spores.tex

\section{Performance evaluation}%
\label{Performance}

\dots