### 2020-03-24 paper/design.tex

%!TEX root = spores.tex

\section{Design of the \name file transfer system}%
\label{design}

% \sonja{will frame and structure (A) squad overlay (prediction) (B)
%   global overlay (RPS) (C) sphinxes (incl.\ sphinx, onion routing in
%   general) (D) SPOR and SPORES (E) encryption (hybrid, payload, header)
%   if not included earlier (F) out-of-band}
% \commentDaniel{I suggest the following structure:
%   (A) SphinxES
%   (B) SPOR (global overlay)
%   (C) SPORES (local overlay).
%   My original idea was to make SphinxES its own section before the design 
%   section instead of a subsection in the design section.}

\include*{spor}
\include*{squad-overlay-design}
%\include*{file-transfer-design}