### 2020-03-27 paper/UCOR.tex

\section{\Acladj*{UC} \acl*{OR}}%
\label{UCOR}

\dots