### 2020-03-27 spores-protocol-paper/abstract.tex

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex spores.tex" ispell-dictionary: "american" -*-

% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?

\acs{SPORES} is a protocol that provides a file transfer service that does not
rely on a central service but provides strong privacy guarantees.

Consider that Alice wants to share a file with Bob. She uploads the file
to \enquote{the Cloud} (someone else's computer) and sends a link to Bob. Bob
fetches the file from the third-party storage. This allows the third
party to infer much information. \Ac{SPORES} changes this.

Alice has several devices (laptop, smartphone, etc.) with varying
online--offline patterns, we call these devices her device squad. The
squad knows which device has which files and each device of the squad
has a model of the other devices' online--offline patterns. Alice's
devices can coordinate the sending of the file. Similarly, Bob also has
a device squad, and his squad can coordinate receiving the file. Instead
of Alice giving Bob a link to the file in the Cloud, Bob gives Alice a
route where the destination is his squad (the \emph{set} of Bob's devices).

At the core of our proposal lies a stateless, probabilistic
onion-routing protocol called \ac{SPOR}. \Ac{SPOR} relies on a set-based
mix-header format based on Sphinx~\cite{Sphinx} (we dubb this extension
SphinxES, for Sphinx Extended for Sets). Given an algorithm for secure
peer sampling, a probabilistic model of node availability, \ac{SPOR} will
populate the layers of SphinxES with sets of devices to reliably deliver
packets from one set of devices to another over a network of mix-nodes
with high churn.

This yields several interesting properties, among others: \ac{SPOR} can
provide connections over a network of unreliable devices, unlike 
Tor's~\cite{Tor} network of dedicated, always online devices; \ac{SPOR} removes 
the weakest-link bandwidth problem present in Tor, sets of nodes at each
layer provides a wider path and thus more bandwidth; at the same time,
\ac{SPOR} still provides privacy guarantees en par with Tor.