### 2020-04-14 paper/spores.tex

\section{\Acf*{SPORES}}%
\label{SPORES}%

We will now extend \ac{SPOR} into \ac{SPORES}.
It is becoming more and more common that people own more than one online device 
(\eg a laptop and a smartphone).
We will utilize this in the construction of \ac{SPORES}.
Thus \ac{SPORES} is different in two aspects:
First, the recipient layer consists of several devices instead of only one.
These devices, the e-squad of Bob, will then coordinate to ensure the relevant 
device gets the message.
We assume there is a protocol which solves this problem.
For instance, \textcite{luxey:cascade} introduce a protocol that allows the 
devices of an e-squad to coordinate user sessions in the face of churn.

Second, we will introduce how to do fragmentation and reassembly for large 
messages.
Consider that the message \(m\) that Alice wants to send to Bob is too large to 
successfully pass every layer of the route.
\Eg each device goes offline and back online every \(t\) seconds, the bandwidth 
is \(w\) bits per second and the message is \(|m|\) bits; if \(|m| > t w\), 
then the message \(m\) can never be delivered as any device will go offline 
before relaying it to the next node.
The natural solution is to split \(m\) into \(n\) parts, \(m_0, \dotsc, 
  m_{n-1}\), where \(|m|/n\leq t w\).
The problem now is that \(m_i\) and \(m_j\) (\(i\neq j\)) might end up on 
different devices.
But that problem is solved by the protocol mentioned above.
This allows us to send large messages, such as videos, more efficiently.

The parameters \(t\) and \(w\) must be empirically approximated and used as 
global configuration parameters for the system.
We run simulations for various values of \(t\) and \(w\) in \cref{Performance}.
In this section we describe a file-transfer protocol for splitting a large 
message, sending the fragments using \ac{SPOR} and reassemble them --- even if 
the fragments are received by different devices of the squad.

\subsection{Fragmentation and reassembly}%
\label{sec:file_transfer}

Alice wants to send the message \(m\) to Bob, \(m\) is \(|m|\) bits long.
Bob creates a reply header and gives it to Alice.
Alice splits \(m\) into \(n\) parts \(m_0, \dotsc, m_{n-1}\).
Alice sends \(|m|\), \(n\), \(\hash(m)\), \(\hash(m_0), \dotsc, 
\hash(m_{n-1})\) to Bob using \ac{SPOR} and Bob's reply header.
(Remember, this means that Bob receives a reply header with the data, so that 
he can later reply to Alice.)

Bob runs \(\CreateReply\) \(n\) times, the node sets passed to \(\CreateReply\) 
are created as in \cref{SPOR}, but all of Bob's devices populate the recipient 
set.
Bob notes the order of the identifiers, \(I\), of each header and sends the 
headers in that order to Alice (using Alice's reply header).

Alice receives the headers and keeps the order of the headers.
For each \(0\leq i < n\), she uses \(\UseReply\) with \(m_i\) as payload to get 
a header \(\hat M_i\).
Then she constructs the final header \(M_i\) by running \(\CreateFwd\) with 
\(\hat M_i\) as payload and the final layer set to \(\{\rdvnode\}\) (to 
indicate to the last node that it is a rendezvous node).
She then sends \(M_i\) along the route.
We note that Alice can process all pieces in parallel as long as she uses the 
headers she got from Bob in the correct order --- \ie \(M_i\) contains \(m_i\) 
as payload.

When the headers are used in the correct order Bob can use the identifiers from 
the headers (\(I\)) to correctly reassemble the fragments despite out-of-order 
delivery.

As above, each payload also contains a reply header for Bob, these can be used 
by Bob to inform Alice at regular time intervals about which chunks have been 
received and which are missing.

%\commentDaniel{I leave the remaining details for someone else to write up.
 % What we need is the details of how Bob's devices syncs and reassembles files.
 % Maybe this is reduced to Sprinkler or something.}